---
title: 'Lab 8: MATH 4753'
author: "Dan Groeneveld"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1

```{r}
getwd()
```

## Task 2

Sample of n=10 from uniform distribution:

```{r}
samp = runif(10,0,5)
samp
```

Mean and variance for uniform distribution where a=0 and b=5:

```{r}
a = 0
b =5
mu = (a +b)/2
var = ((b-a)^2)/12
mu
var
```

Mean and variance from sample:

```{r}
xbar = mean(samp)
ssqr = var(samp)
xbar
ssqr
```

Using the above theory:

$$ \mu_T = \frac{n(b+a)}{2}$$

$$ \sigma_T =  \frac{n(b-a)^2}{12}$$

$$\mu_\bar{Y} = \frac{b+a}{2}$$

$$\sigma_\bar{Y} =\frac{(b-a)^2}{12n}$$

Explaining parts of the function myclt:

A - This part creates a sample of n from a uniform function with a and b (equal to 0 and 5 in the example) and then iterates the number of times given in the value iter. This object is then stored in the variable y.

B - This part turns the object y from before into a matrix with the number of rows equaling the sample size n and the number of columns equaling the number iterations iter. This matrix is stored in the variable data.

C - This part applies a sum to the columns of the matrix of data and stores these values in the variable sm.

D - This line calls the function myclt with values of n=10 and iter=1000 and then places the functions values in the object w.

```{r}
myclt=function(n,iter,a=0,b=5){
y=runif(n*iter,a,b)
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
sm=apply(data,2,sum)
h=hist(sm,plot=FALSE)
hist(sm,col=rainbow(length(h$mids)),freq=FALSE,main="Distribution of the sum of uniforms")
curve(dnorm(x,mean=n*(a+b)/2,sd=sqrt(n*(b-a)^2/12)),add=TRUE,lwd=2,col="Blue")
sm
}
w=myclt(n=10,iter=10000)
```

```{r}
wbar = mean(w)
ssqrw = var(w)
wbar
ssqrw
```

Histogram of sample means and calculated values for w:

```{r}
mycltmeans=function(n,iter,a=0,b=5){
y=runif(n*iter,a,b)
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
means=apply(data,2,mean)
h=hist(means,plot=FALSE)
hist(means,col=rainbow(length(h$mids)),freq=FALSE,main="Distribution of the sample means")
curve(dnorm(x,mean=(a+b)/2,sd=sqrt((b-a)^2/12/n)),add=TRUE,lwd=2,col="Blue")
means
}
w=mycltmeans(n=10,iter=10000)
mean(w)
var(w)
```

## Task 3

Examining the function:

```{r}
mycltu=function(n,iter,a=0,b=10){
## r-random sample from the uniform
y=runif(n*iter,a,b)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax
## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), main=paste("Histogram of sample mean",
"\n", "sample size= ",n,sep=""),xlab="Sample mean")
## add a density curve made from the sample distribution
lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
## Add the density from which the samples were taken
curve(dunif(x,a,b),add=TRUE,lwd=4)

}


```

The 2 in the apply function means that the mean of each column of data is taken.

There are 20 * 100,000 terms in w when mycltu is called with n=20 and iter =100000.

sd takes the form as shown in the equation because this is what the standard deviation of a sample taken from a uniform function will approach as n tends toward infinity. 

Plots:

```{r}
mycltu(n=1,iter=10000)
mycltu(n=2,iter=10000)
mycltu(n=3,iter=10000)
mycltu(n=5,iter=10000)
mycltu(n=10,iter=10000)
mycltu(n=30,iter=10000)
```

I conclude that with increasing n (the sample size) the distribution of the sample means is better approximated by a normal function with mean: 
$$\mu = \frac{b+a}{2}$$

and standard deviation: 
$$\sigma =\frac{(b-a)^2}{12n}$$ 



## Task 4

Samples from the binomial distribution:

```{r}
mycltb=function(n,iter,p=0.5,...){

## r-random sample from the Binomial
y=rbinom(n*iter,size=n,prob=p)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Now we can make the histogram
## freq=FALSE means take a density
hist(w,freq=FALSE,  ylim=c(0,ymax),
main=paste("Histogram of sample mean","\n", "sample size= ",n,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=n*p,sd=sqrt(p*(1-p))),add=TRUE,col="Red",lty=2,lwd=3) 

}



```

### p = 0.3

```{r}
mycltb(n=4,iter=10000,p=0.3)
mycltb(n=5,iter=10000,p=0.3)
mycltb(n=10,iter=10000,p=0.3)
mycltb(n=20,iter=10000,p=0.3)
```

### p = 0.7

```{r}
mycltb(n=4,iter=10000,p=0.7)
mycltb(n=5,iter=10000,p=0.7)
mycltb(n=10,iter=10000,p=0.7)
mycltb(n=20,iter=10000,p=0.7)
```

### p = 0.5

```{r}
mycltb(n=4,iter=10000,p=0.5)
mycltb(n=5,iter=10000,p=0.5)
mycltb(n=10,iter=10000,p=0.5)
mycltb(n=20,iter=10000,p=0.5)
```

I conclude that regardless of the probability, the distribution of sample means taken from a binomial distribution approaches a normal distribution with increasing sample size n. The probability only changes where the normal distribution is centered around.

## Task 5

Samples from a poisson distribution:

```{r}
mycltp=function(n,iter,lambda=10,...){

## r-random sample from the Poisson
y=rpois(n*iter,lambda=lambda)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Make a suitable layout for graphing
layout(matrix(c(1,1,2,3),nr=2,nc=2, byrow=TRUE))

## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), col=rainbow(max(w)),
main=paste("Histogram of sample mean","\n", "sample size= ",n," iter=",iter," lambda=",lambda,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=lambda,sd=sqrt(lambda/n)),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve

# Now make a new plot
# Since y is discrete we should use a barplot
barplot(table(y)/(n*iter),col=rainbow(max(y)), main="Barplot of sampled y", ylab ="Rel. Freq",xlab="y" )
x=0:max(y)
plot(x,dpois(x,lambda=lambda),type="h",lwd=5,col=rainbow(max(y)),
main="Probability function for Poisson", ylab="Probability",xlab="y")
}



```

### lambda = 4

```{r}
mycltp(n=2,iter=10000, lambda = 4)
mycltp(n=3,iter=10000, lambda = 4)
mycltp(n=5,iter=10000, lambda = 4)
mycltp(n=10,iter=10000, lambda = 4)
mycltp(n=20,iter=10000, lambda = 4)
```

### lambda = 10

```{r}
mycltp(n=2,iter=10000, lambda = 10)
mycltp(n=3,iter=10000, lambda = 10)
mycltp(n=5,iter=10000, lambda = 10)
mycltp(n=10,iter=10000, lambda = 10)
mycltp(n=20,iter=10000, lambda = 10)
```

## Task 6

Installing the package (this was done and then blocked out to avoid screeds of output)

```{r}
#devtools::install(build_vignettes=TRUE)
```

Calling the function:

```{r}
w = MATH4753GROE002::mycltdangroeneveld(n=10, iter =10000)
```








