---
title: 'Lab 5: MATH 4753'
author: "Dan Groeneveld"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Task 1


```{r}
getwd()
```


## Task 2

```{r}
library(MATH4753GROE002)
mybin(iter=100, n=10, p=0.7)
mybin(iter=200, n=10, p=0.7)
mybin(iter=500, n=10, p=0.7)
mybin(iter=1000, n=10, p=0.7)
tabsim <- mybin(iter=10000, n=10, p=0.7)

```

The table for 10,000 iterations:

```{r}
tabsim
```

```{r}
dbin.tab = round(dbinom(0:10, size = 10, prob = 0.7), 4)
names(dbin.tab) = 0:10
dbin.tab
```

```{r}
tab <- round(dbinom(0:10, size = 10, prob = 0.7), 4)
tab - tabsim
```

Yes, these values are approximately correct as the differences are all relatively small.



## Task 3

```{r}

marb <- c("1","1","1","1","1","1","1","1","1","1","1","1","0","0","0","0","0","0","0","0")
p <- c(3/5,3/5,3/5,3/5,3/5,3/5,3/5,3/5,3/5,3/5,3/5,3/5,2/5,2/5,2/5,2/5,2/5,2/5,2/5,2/5)
sample(marb, size =5, prob=p, replace = FALSE)
sample(marb, size =5, prob=p, replace =TRUE)  
```

```{r}
myhyper=function(iter=100,N=20,r=12,n=5){
# make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
#Make a vector to hold the number of successes over the trials
succ=c()
for( i in 1:iter){
#Fill each column with a new sample
sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}
#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))
#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
succ.tab/iter
}

myhyper(iter=100,n=5, N=20,r=12)
myhyper(iter=200,n=5, N=20,r=12)
myhyper(iter=500,n=5, N=20,r=12)
myhyper(iter=1000,n=5, N=20,r=12)
esthyp <- myhyper(iter=10000,n=5, N=20,r=12)

```

The table for 10,000 iterations:

```{r}
esthyp
```


```{r}
acthyp <- dhyper(x=0:5, m=12, n= 8, k=5)
acthyp
esthyp - acthyp
```

Yes, these values are approximately correct as the differences are all relatively small.

## Task 4

I think the function mysample generates a set of random data with 10 possibilities and n data points and then plots a sample drawing assuming the likelihood of drawing any one of 10 possibilities is the same. It iterates this process according to the value of iter and plots these in R at intervals given by time (in seconds).

```{r}
mysample=function(n, iter=10,time=0.5){
for( i in 1:iter){
#make a sample
s=sample(1:10,n,replace=TRUE)
# turn the sample into a factor
sf=factor(s,levels=1:10)
#make a barplot
barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
ylim=c(0,0.2)
)

#release the table
Sys.sleep(time)
}
}

#mysample(n=1000, iter=30, time = 1)

```

I see 30 plots constructed all showing (if roughly) even distributions for the 10 possible outcomes. The plots appear at 1 second intervals in R. The approximately even distribution of possibilities show that each possibility had an equal likelihood.

```{r}
mysample(n=1000, iter=1, time = 1)
```


## Task 5

```{r}
choose(8,4)
```

```{r}
1 - ppois(4, 2) 
```

```{r}
dnbinom(7,3,0.4) 
```

```{r}
pbinom(8,15,0.4)
```

## Task 6

```{r}
MATH4753GROE002::dansample(n=10, iter = 5, time = 2)
```



