---
title: "Lab1"
author: "Dan Groeneveld"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1


The working directory

```{r}
getwd()
```
## Task 2

Read in DDT data

```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```


## Task 3

  a. The qualitative variables are RIVER and SPECIES.
  b. The quantitative variables are MILE, LENGTH, WEIGHT, and DDT.
  
```{r}
sp <- ddt$SPECIES
tab <- table(sp)
barplot(tab,col = 1:3)
```
 
  c. There are three species in the ddt data set.
  d.
  
```{r}
bigfish <- with(ddt, ddt[WEIGHT>800 & SPECIES == "LMBASS",])
head(bigfish)
```

  e.
  
```{r}
bigddt <- with(ddt, ddt[RIVER == "SCM" & DDT>4.0,])
head(bigddt)
```

## Clicker Questions

  a. The mean length of fish in the ddt data frame is 43.
```{r}
len=with(ddt,LENGTH)
mean(len)
```
  
  b. The standard deviation of fish weight in the ddt data frame is 377.
```{r}
stdwgt=with(ddt,WEIGHT)
sd(stdwgt)
```

  c. No, the plot pictured on the lab assignment has the reversed axis scales of the correct plot from the ddt data frame (pictured below).
```{r}
library(ggplot2)
s = ggplot(ddt, aes(x = WEIGHT, y = LENGTH))
 s = s + geom_point(aes(color= SPECIES)) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Length vs. Weight",
       caption="Source: DDT.csv",
       x="Weight",
       y="Length")

s
```
  
  d. The last value is 1.00.
```{r}
v=1:20
v/20
```
  1.00
  
## Task 4

```{r}
rtab <- with(ddt, table(RIVER))
rtab
barplot(rtab, col = 1:4)
```


```{r}
rivsp=with(ddt,table(RIVER,SPECIES))
rivsp
barplot(rivsp,beside=TRUE,col=rainbow(4))

```


## Task 5

```{r}
pie(tab, col=1:4)
```

```{r}
pie(rtab, col=1:4)
```

## Task 6

```{r}
layout(matrix(c(1,2,3),nr=1,nc=3))# 1 row 3 cols

with(ddt,boxplot(DDT,ylab="DDT",col="Red",notch=TRUE))
with(ddt,boxplot(WEIGHT,ylab="WEIGHT",col="Green",notch=TRUE))
with(ddt,boxplot(LENGTH,ylab="LENGTH",col="Blue",notch=TRUE))

```

## Task 7

```{r}
fishcol=with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))

#River col
rivcol=with(ddt, ifelse(RIVER=="FCM","Red",
                        ifelse(RIVER=="LCM","Blue",
                               ifelse(RIVER=="SCM","Green","Black"))))




#Given the combinations of all levels of SPECIES and RIVER = RIVER*SPECIES
coplot(LENGTH~WEIGHT|RIVER,data=ddt,col=rivcol)
coplot(DDT~WEIGHT|SPECIES,data=ddt,col=fishcol)
```

## Task 8

```{r}
library(ggplot2)
#library(ggpubr)
#library(ggpmisc)
g = ggplot(ddt, aes(x = LENGTH, y = WEIGHT )) + 
  geom_point(aes( shape = SPECIES))


lbyr <- c(12,13,14)*100

 b = ggplot(ddt, aes(x = SPECIES, y = WEIGHT))
 b = b + geom_boxplot(aes(fill= RIVER)) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Dan Groeneveld", 
       caption="Source: DDT.csv",
       x="Species",
       y="Weight")

b

 v = ggplot(ddt, aes(x = RIVER, y = LENGTH))
 v = v + geom_violin(aes(fill= SPECIES)) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Dan Groeneveld", 
       caption="Source: DDT.csv",
       x="River",
       y="Length")

v

 s = ggplot(ddt, aes(x = WEIGHT, y = LENGTH))
 s = s + geom_point(aes(color= SPECIES)) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Dan Groeneveld", 
       caption="Source: DDT.csv",
       x="Weight",
       y="Length")

s

```



