---
title: 'Lab 11: MATH 4753'
author: "Dan Groeneveld"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
d=c(5.0581, 4.9707, 5.0893, 4.9334, 4.9777, 5.0285, 4.8555, 4.9565, 
4.9769, 4.9722, 4.999, 4.9925, 4.9686, 5.0662, 4.9239, 4.9781, 
5.0485, 5.0014, 4.9957, 5.0195, 5.0118, 4.9928, 5.0361, 5.0185, 
4.9879)

```

### 95% CI for mu

```{r}
t=qt(0.975,24)
ci=c()
ci[1]=mean(d)-t*sd(d)/sqrt(25)
ci[2]=mean(d)+t*sd(d)/sqrt(25)
ci
```

### 90% CI for mu

```{r}
t=qt(0.95,24)
ci=c()
ci[1]=mean(d)-t*sd(d)/sqrt(25)
ci[2]=mean(d)+t*sd(d)/sqrt(25)
ci
```

### 80% CI for mu

```{r}
t=qt(0.90,24)
ci=c()
ci[1]=mean(d)-t*sd(d)/sqrt(25)
ci[2]=mean(d)+t*sd(d)/sqrt(25)
ci
```

### 50% CI for mu

```{r}
t=qt(0.75,24)
ci=c()
ci[1]=mean(d)-t*sd(d)/sqrt(25)
ci[2]=mean(d)+t*sd(d)/sqrt(25)
ci
```

### 80% CI for mu using t.test()

```{r}
obj=t.test(d,conf.level=0.80)
obj$conf.int
```

### 95% CI for sigma squared

```{r}
chialpha2 = qchisq(1-.05/2, 25-1)
chiminalpha2 = qchisq(.05/2, 25-1)
ci= c()
ci[1] = ((25-1) * var(d))/chialpha2
ci[2] = ((25-1) * var(d))/chiminalpha2
ci
```

### 90% CI for sigma squared

```{r}
chialpha2 = qchisq(1-.10/2, 25-1)
chiminalpha2 = qchisq(.10/2, 25-1)
ci= c()
ci[1] = ((25-1) * var(d))/chialpha2
ci[2] = ((25-1) * var(d))/chiminalpha2
ci
```

### 80% CI for sigma squared

```{r}
chialpha2 = qchisq(1-.20/2, 25-1)
chiminalpha2 = qchisq(.20/2, 25-1)
ci= c()
ci[1] = ((25-1) * var(d))/chialpha2
ci[2] = ((25-1) * var(d))/chiminalpha2
ci
```

### 50% CI for sigma squared

```{r}
chialpha2 = qchisq(1-.50/2, 25-1)
chiminalpha2 = qchisq(.50/2, 25-1)
ci= c()
ci[1] = ((25-1) * var(d))/chialpha2
ci[2] = ((25-1) * var(d))/chiminalpha2
ci
```

## Task 3

```{r}
blue=c(21.65, 17.48, 20.1, 21.57, 14.82, 19.17, 21.08, 18.23, 22.93, 
15.66, 20.89, 21.66, 18.5, 20.59, 18.63, 18.91, 19.53, 17.7, 16.5, 19.03)
snapper=c(31.65, 27.48, 30.1, 31.57, 24.82, 29.17, 31.08, 28.23, 32.93, 
25.66, 30.89, 31.66, 28.5, 30.59, 28.63)

```

### 95% CI for musnapper - mublue

```{r}
n1=length(snapper)
n2=length(blue)
spsq=((n1-1)*var(snapper)+(n2-1)*var(blue))/(n1+n2-2)
t=qt(0.975,n1+n2-2)
ci=c()
ci[1]=mean(snapper)-mean(blue)-t*sqrt(spsq*(1/n1+1/n2)) 
ci[2]=mean(snapper)-mean(blue)+t*sqrt(spsq*(1/n1+1/n2))
ci
```

This means that for 95% of the pairs of caught blue and snapper fish, the snapper will be between 8.76 and 11.84 in longer than the blue.

### 95% CI using t.test()

```{r}
t.test(snapper,blue,conf.level=0.95,var.equal=TRUE)$conf.int
```

### 85% CI using t.test()

```{r}
t.test(snapper,blue,conf.level=0.85,var.equal=TRUE)$conf.int
```

### 75% CI using t.test()

```{r}
t.test(snapper,blue,conf.level=0.75,var.equal=TRUE)$conf.int
```

### 25% CI using t.test()

```{r}
t.test(snapper,blue,conf.level=0.25,var.equal=TRUE)$conf.int
```

The interval decreases in size (the upper and lower bounds are closer together) as the confidence decreases.

## Task 4

```{r}
Exam1=c(40.98, 59.36, 46.69, 41.8, 61.63, 65.31, 62.96, 60.21, 56.89, 
78.41, 53.44, 75.2, 60.54, 52.43, 41.41, 70.79, 73.55, 55.65, 
61.43, 63.84, 58.07, 53.79, 54.45, 67.18, 44.46)

Exam2=c(50.22, 66.19, 58.75, 51.88, 66.61, 70.86, 74.25, 70.23, 69.55, 
87.18, 63.62, 81.7, 70.5, 66.02, 51.35, 80.92, 85.65, 65.44, 
74.37, 75.28, 67.86, 59.92, 64.42, 73.57, 57.15)

```

### 95% CI for the mean difference

```{r}
exdif = Exam1 - Exam2
muexdif = mean(exdif)
sdexdif = sd(exdif)
n1 = length(Exam1)
ci = c()
t=qt(0.975,n1-1)
ci[1] = muexdif - t * sdexdif/sqrt(n1)
ci[2] = muexdif + t * sdexdif/sqrt(n1)
ci
```

This means that for 95% of students, they did between 8.71 and 10.73 points worse on the first exam than on the second.

### 90% CI using t.test()

```{r}
t.test(exdif, conf.level = .90)$conf.int
```

### 80% CI using t.test()

```{r}
t.test(exdif, conf.level = .80)$conf.int
```

### 70% CI using t.test()

```{r}
t.test(exdif, conf.level = .70)$conf.int
```

### 60% CI using t.test()

```{r}
t.test(exdif, conf.level = .60)$conf.int
```

### 10% CI using t.test()

```{r}
t.test(exdif, conf.level = .10)$conf.int
```

## Task 5

```{r}
NZBIRDS = read.csv('NZBIRDS.csv')
tab = with(NZBIRDS, table(Extinct,Flight))
addmargins(tab)
```

```{r}
z=1.96 #Taken from .pdf provided on assignment page
n1p1 = 38 * .5526
n1q1 = 38 * .4474
n2p2 = 78 * 0.0897
n2q2 = 78 * .9103 # all these values are large enough to use normal approximation
```

### 95% ci for difference in proportion of flightless birds for extinct and non-extinct species

```{r}
ci = c()
ci[1] = .5526 - 0.0897 - z*sqrt((.5526*.4474)/38 + (.0897*.9103)/78)
ci[2] = .5526 - 0.0897 + z*sqrt((.5526*.4474)/38 + (.0897*.9103)/78)
ci
```

## Task 6

### 95% CI for ratio of variances

```{r}
set.seed(35)
sam1=rnorm(25,mean=10,sd=5)
set.seed(45)
sam2=rnorm(34,mean=40,sd=8)
a=1-.05/2
ci=c()
ci[1]=var(sam1)/(var(sam2)*((qf(a, df1=24,df2=33))))
ci[2]=var(sam1)/var(sam2)*((qf(a, df1=33,df2=24)))
ci
```

### 80% CI using var.test()

```{r}
var.test(sam1,sam2, conf.level =.80)
```

### 70% CI using var.test()

```{r}
var.test(sam1,sam2, conf.level =.70)
```

### 60% CI using var.test()

```{r}
var.test(sam1,sam2, conf.level =.60)
```

### 50% CI using var.test()

```{r}
var.test(sam1,sam2, conf.level =.50)
```

## Task 7 

Using the function I created

```{r}
set.seed(23)
x = rnorm(30,mean=10,sd=12)
MATH4753GROE002::myci(d=x, n=30, a=0.05)
```










