---
title: "Assignment 1: MATH 4753"
author: "Dan Groeneveld"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

14/15 questions completed.

## 1-Grading

My final grade for the course will be calculated as the following:
```{r}
read.csv("Gradingscheme.csv")
```

The right most column sums to 100% as expected.
As in the class are 90s, Bs are 80s, Cs are 60s and 70s, Ds are 50s, and below 50 is failing. There is NO curving for this course!

## 2-Coplot

```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
m=with(ddt, as.numeric(factor(MILE)))
length(unique(m))
```

### a. 

```{r}
coplot(LENGTH~WEIGHT|RIVER*SPECIES,data=ddt,col=m, pch=21)
```

### b.

The lower left 3 conditional plots show that only one species of fish (C Catfish) were caught in each of the rivers (FCM, LCM, SCM). 

### c.

Line A turns the variable mile into a numeric value.

### d.

Line B shows the number of unique miles in the dataset.

### e.

The top 6 plots are empty becuase no fish of the species LMBASS or SMBUFFALO were caught in the rivers FCM, LCM, or SCM.

### f.

```{r}
fcmddt <- ddt[ddt$RIVER=="FCM" & ddt$SPECIES=="CCATFISH",]$DDT
mean(fcmddt)
```

The mean value of ddt found CCATFISH from the FCM river is 45.

## 3-National Bridge Inventory

### a.

Quantitative

### b. 

Quantitative

### c. 

Qualitative

### d. 

Quantitative

### e. 

Qualitative

### f.

Quantitative

### g.

Qualitative
  
## 4-Types of Random Sampling

### a.

The 4 types of random sampling are simple random sampling, stratified random sampling, cluster sampling, and systematic sampling.

### b.

Simple random sampling is choosing at random a predetermined number of experimental units from a population. Stratified random sampling is used when a population can be divided into two or more groups. Random samples are then taken from each group before being combined to form the overall sample of the population. Cluster sampling involves taking a complete sample of certain groups within a population and then combining these to form a sample of the total population. This is usually used when it would be difficult or costly to take small samples from each group within a population. Systematic sampling involves sampling every kth unit of a population to obtain a representative sample. 

## 5-MTBE Well Problem

```{r}
mtbe=read.csv("MTBE.csv", header=TRUE) 
head(mtbe) 
dim(mtbe) 
ind=sample(1:223,5,replace=FALSE) # random indices
mtbe[ind,]
```

The wells in my sample are 118, 23, 13, 148, and 131.


### a. 

#### i.
```{r}
mtbeo=na.omit(mtbe)
head(mtbeo)
```
   
#### ii.
```{r}
depth=mtbeo[mtbeo$Aquifier=="Bedrock",]$Depth
mean(depth)
sd(depth)
```
  
The standard deviation of well depth of wells with Bedrock as the aquifier is 56.4536.

## 6-Earthquake Problem

```{r}
eq=read.csv("EARTHQUAKE.csv", header=TRUE) 
head(eq) 
dim(eq)
v = sample(1:2929,10,replace=FALSE)
eq[v,]
```

### a.

#### i.

```{r}
plot(ts(eq$MAG))
```

#### ii.
```{r}
mag = eq$MAGNITUDE
median(mag)
```
The median value for magnitude of the earthquake is 2.


## 7-MS Statistics In Action

### a. 

The data collection method is a designed experiment with a stratified sample. The different locations where specimens of fish were collected represent the strata for this study.

### b.

The population for this study is not the sum of all the fish specimens that were obtained but rather the total number of fish in the Tennessee River and its various tributaries.

### c.

The names of all the qualitative variables are "MILE", "RIVER", and "SPECIES".

## 8-Do Social Robots Walk or Roll?

```{r}
freq=c(15,8,63,20)
RL=c("None","Both","LegsO","WheelsO")
l=rep(RL,freq)
lt = table(l)
barplot(lt)
```


### a. 

A barplot is used to describe the data.

### b. 

The variable measured was presence of wheels and/or legs in each of the 103 robot designs.

### c. 

Legs only is the robot design that is currently most used.

### d. 

The class frequencies for each category are as follows: None: 0.1415, Both: 0.0755, Legs ONLY: 0.5943, Wheels ONLY: 0.1887.

### e. 

Pareto Diagram
  
```{r}
pareto<-function(x,mn="Pareto barplot",...){  # x is a vector
x.tab=table(x)
xx.tab=sort(x.tab, decreasing=TRUE,index.return=FALSE)
cumsum(as.vector(xx.tab))->cs
length(x.tab)->lenx
bp<-barplot(xx.tab,ylim=c(0,max(cs)),las=2)
lb<-seq(0,cs[lenx],l=11)
axis(side=4,at=lb,labels=paste(seq(0,100,length=11),"%",sep=""),las=1,line=-1,col="Blue",col.axis="Red")
for(i in 1:(lenx-1)){
segments(bp[i],cs[i],bp[i+1],cs[i+1],col=i,lwd=2)
}
title(main=mn,...)

}

pareto(l)
```

## 9-Microsoft Program Security Issues

### a.

```{r}
freq1=c(32,6,12)
products=c("Windows","Explorer","Office")
mp=rep(products,freq1)
mpt = table(mp)
pie(mpt, col = 2:4)
```

Explorer was the Microsoft product with the lowest proportion of security issues in 2012.

### b.

```{r}
freq2=c(6,8,22,3,11)
repercuss=c("DoS","ID","RCE", "S", "PE")
mr=rep(repercuss,freq2)
pareto(mr)
```

Based on the graph, I would advise Microsoft to focus on remote code execution.

## 10-Software Defects

```{r}
swd=read.csv("SWDEFECTS.csv", header=TRUE) 
head(swd) 
library(plotrix)
tab=table(swd$defect)
rtab=tab/sum(tab)
round(rtab,2)
pie3D(rtab,labels=list("OK","Defective"),main="pie plot of SWD")
```

It is relatively unlikely that a given piece of software code will be defective, given that a clear majority of software code is OK. That being said, a sizable portion of code is still defective.
  
  
## 11-Process Voltage Readings

### a.

```{r}
volt=read.csv("Oldlocationtable.csv", header=TRUE)
volt
barplot(volt$Relative.Frequency, space = 0, xlab = "Class Interval", ylab = "Relative Frequency", main = "Old Voltage Histogram", col = 2:11)
```

### b.

```{r}
voltage = read.csv("VOLTAGE.csv", header = TRUE)
head(voltage)
oldv = voltage[voltage$LOCATION == "OLD",]$VOLTAGE
stem(oldv)
```

The stem-and-leaf display is more informative about where most of the voltage readings lie as it clearly shows the cluster from 9.6-10.3

### c.

```{r}
voltage.df = data.frame(voltage)
new<-subset(voltage.df,subset=LOCATION=="NEW")
new$VOLTAGE->vtn
vtn
max(vtn)
min(vtn)
lept<-min(vtn)-0.05
rept<-max(vtn)+0.05
rnge<-rept-lept
inc<-rnge/9
inc
seq(lept, rept,by=inc)->cl
cl
cvtn<-cut(vtn,breaks=cl)
new.tab=table(cvtn)
barplot(new.tab,space=0,main="Frequency Histogram(NEW)",las=2)
hist(vtn,nclass=10)
```

### d.

It appears that the manufacturing process cannot be established locally, the new process is not as good as the old process, as there are more readings below 9.2 volts.

### e.

```{r}
mean(oldv)
median(oldv)
mode(oldv)
mean(vtn)
median(vtn)
mode(vtn)
```
The mean of the old voltage dataset is 9.8037.
The median of the old voltage dataset is 9.975. 
The mode is not meaningful in this situation.
The mean of the new voltage dataset is 9.4223.
The median of the new voltage dataset is 9.455.
The mode is not meaninful in this situation.



## 12-Surface Roughness of Pipe


```{r}
rp=read.csv("ROUGHPIPE.csv", header=TRUE)
rough=rp$ROUGH
hist(rough, col = 2:12)
mean(rough)
sd(rough) * 2
```

The interval from 0.8332 to 2.9288 micrometers will likely contain 95% of all coated pipe roughness measurements by the empirical rule.


## 13-Mongolian Desert Ants

### a.

```{r}
ga=read.csv("GOBIANTS.csv", header=TRUE)
head(ga)
as = ga$AntSpecies
mean(as)
median(as)
mode(as)
```

The mean is 12.8182
The median is 5
The modes are 4 and 5

### b.

I would recommend using the median as the measure of central tendency to describe the center of the number of ant species distribution because the mean is skewed by two outliers, having 49 and 52 ant species respectively. The mode is also useful, but because there are two modes, and one of the modes is the median, I think the median does a better job at representing the central tendency.

### c.

```{r}
dstpc=ga[ga$Region=="Dry Steppe",]$PlantCov
mean(dstpc)
median(dstpc)
mode(dstpc)
```

The mean is 40.4%
The median is 40%
The mode is 40%

### d.

```{r}
gdtpc=ga[ga$Region=="Gobi Desert",]$PlantCov
mean(gdtpc)
median(gdtpc)
mode(gdtpc)
```

The mean is 28%
The median is 26%
The mode is 30%

### e.

Yes, the center of total plant cover percentage distribution appears to be different at the two regions. In the dry steppe region, the center of total plant cover percentage distribution is around 40% while it's around 26-30% in the gobi desert region.


## 14-Speed of Light from Galaxies

### a.

```{r}
library(lattice)
g2 = read.csv("GALAXY2.csv", header = TRUE)
head(g2)
vel=g2$VELOCITY
dotplot(vel, main = "Dotplot of Velocities")
```

### b.

Yes, there is evidence to support the double cluster theory. In the dotplot above, there are two distinct groupings of velocities, one centered around roughly 19500 km/s and the other centered roughly around 22500 km/s.

### c.

```{r}
A1775A = g2[g2$VELOCITY < 21000,]
A1775B = g2[g2$VELOCITY > 21000,]
mean(A1775A)
sd(A1775A)
mean(A1775B)
sd(A1775B)
```

The mean velocity for galaxy cluster A1775A is 19462.24 km/s and the standard deviation is 532.2868 km/s.
The mean velocity for galaxy cluster A1775B is 22838.47 km/s and the standard deviation is 560.9767 km/s.

### d.

A galaxy of velocity 20,000 km/s is likely to belong to cluster A1775A because this velocity falls within roughly one standard deviation of the mean galaxy velocity for cluster A1775A.


## 15-ggplot


```{r}
library(ggplot2)
ddt = read.csv("DDT.csv", header = TRUE)

 b = ggplot(ddt, aes(x = RIVER, y = LENGTH))
 b = b + geom_boxplot(aes(fill= SPECIES)) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Dan Groeneveld", 
       caption="Source: DDT.csv",
       x="RIVER",
       y="LENGTH")
b

```




