---
title: 'Lab 6: MATH4753'
author: "Dan Groeneveld"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Task 1

Working directory.

```{r}
getwd()
```

## Task 2

Plotting normal distributions.

```{r}
#Setting up the layout to show 4 plots

layout(matrix(1:4, nr=2,nc=2))

#layout.show(4)

#Making the normal distribution functions

curve(dnorm(x, mean=10,sd=4),xlim=c(-10,30))

curve(dnorm(x, mean=10,sd=2),xlim=c(0,20))

curve(dnorm(x, mean=5,sd=10),xlim=c(-20,30))

curve(dnorm(x, mean=5,sd=0.5),xlim=c(2,8))
```

Plotting probabilities for normal distributions

```{r}

# First curve
curve(dnorm(x, mean=0,sd=1),xlim=c(-5,5))

# Find the area greater than 2

# x values corresponding to the x - cords of points on the curve
xcurve1=seq(2,10,length=1000)

# Y values corresponding t0 the x values
ycurve1=dnorm(xcurve1,mean=0,sd=1)

# Fill in the polygon with the given vertices
polygon(c(2,xcurve1,10),c(0,ycurve1,0),col="Red")

# Put in the text with the appropriate area

# Area
prob1=1 - pnorm(2,mean=0,sd=1)
prob1=round(prob1,4)

# Click to paste the text onto the graph

text(2,0.1, paste("Area = ", prob1, sep=""))

# Second curve
curve(dnorm(x, mean=4,sd=2),xlim=c(-2,10))

# Find the area between x=1 and 5

# x values corresponding to the x - cords of points on the curve
xcurve2=seq(1,5,length=1000)

# Y values corresponding t0 the x values
ycurve2=dnorm(xcurve2,mean=4,sd=2)

# Fill in the polygon with the given vertices
polygon(c(1,xcurve2,5),c(0,ycurve2,0),col="Red")

# Put in the text with the appropriate area

# Area
prob2=pnorm(5,mean=4,sd=2)-pnorm(1,mean=4,sd=2)
prob2=round(prob2,4)

# Click to paste the text onto the graph

text(3,0.05, paste("Area = ", prob2, sep=""))

# third curve
curve(dnorm(x, mean=10,sd=4),xlim=c(-6,28))

# Find the area between x=-6 and 10

# x values corresponding to the x - cords of points on the curve
xcurve3=seq(-6,10,length=1000)

# Y values corresponding t0 the x values
ycurve3=dnorm(xcurve3,mean=10,sd=4)

# Fill in the polygon with the given vertices
polygon(c(-6,xcurve3,10),c(0,ycurve3,0),col="Red")

# Put in the text with the appropriate area

# Area
prob3=pnorm(10,mean=10,sd=4)
prob3=round(prob3,4)

# Click to paste the text onto the graph

text(8,0.05, paste("Area = ", prob3, sep=""))

# third curve
curve(dnorm(x, mean=-2,sd=0.5),xlim=c(-4,0))

# Find the area between x=-6 and 10

# x values corresponding to the x - cords of points on the curve
xcurve4=seq(-3,-2,length=1000)

# Y values corresponding t0 the x values
ycurve4=dnorm(xcurve4,mean=-2,sd=0.5)

# Fill in the polygon with the given vertices
polygon(c(-3,xcurve4,-2),c(0,ycurve4,0),col="Red")

# Put in the text with the appropriate area

# Area
prob4=pnorm(-2,mean=-2,sd=0.5) - pnorm(-3,mean=-2,sd=0.5)
prob4=round(prob4,4)

# Click to paste the text onto the graph

text(-2.5,0.4, paste("Area = ", prob4, sep=""))

```

## Task 3


Plotting the gamma distributions.

```{r}
curve(dgamma(x,shape=1,scale=1),xlim=c(0,10),ylim=c(0,1),col="Red",lwd=2,
ylab="Gamma density", main="Beta=1")
text(0.9, 0.8,paste("alpha=",1))
curve(dgamma(x,shape=3,scale=1),xlim=c(0,10),ylim=c(0,1),add=TRUE,lwd=2)
text(2, 0.3,paste("alpha=",3))
curve(dgamma(x,shape=5,scale=1),xlim=c(0,10),ylim=c(0,1),add=TRUE, col="Blue",lwd=2)
text(5,0.22,paste("alpha=",5))
```

Plotting probabilities for the gamma distributions.

```{r}
# First curve
curve(dgamma(x, shape=3,scale=2),xlim=c(0,15))

# Find the area between 2 and 5

# x values corresponding to the x - cords of points on the curve
xcurveg1=seq(2,5,length=1000)

# Y values corresponding t0 the x values
ycurveg1=dgamma(xcurveg1,shape=3,scale=2)

# Fill in the polygon with the given vertices
polygon(c(2,xcurveg1,5),c(0,ycurveg1,0),col="Red")

# Put in the text with the appropriate area

# Area
probg1=pgamma(5,shape=3,scale=2) - pgamma(2,shape=3,scale=2)
probg1=round(probg1,4)

# Click to paste the text onto the graph

text(3.5,0.06, paste("Area = ", probg1, sep=""))
```

```{r}
# Second curve
curve(dgamma(x, shape=6,scale=3),xlim=c(0,40))

# Find the area between 1 and 4

# x values corresponding to the x - cords of points on the curve
xcurveg2=seq(1,4,length=1000)

# Y values corresponding t0 the x values
ycurveg2=dgamma(xcurveg2,shape=6,scale=3)

# Fill in the polygon with the given vertices
polygon(c(1,xcurveg2,4),c(0,ycurveg2,0),col="Red")

# Put in the text with the appropriate area

# Area
probg2=pgamma(4,shape=6,scale=3) - pgamma(1,shape=6,scale=3)
probg2=round(probg2,4)

# Click to paste the text onto the graph

text(2.5,0.01, paste("Area = ", probg2, sep=""))
```

```{r}
# Third curve
curve(dgamma(x, shape=2,scale=4),xlim=c(0,25))

# Find the area between 3 and 6

# x values corresponding to the x - cords of points on the curve
xcurveg3=seq(3,6,length=1000)

# Y values corresponding t0 the x values
ycurveg3=dgamma(xcurveg3,shape=2,scale=4)

# Fill in the polygon with the given vertices
polygon(c(3,xcurveg3,6),c(0,ycurveg3,0),col="Red")

# Put in the text with the appropriate area

# Area
probg3=pgamma(6,shape=2,scale=4) - pgamma(3,shape=2,scale=4)
probg3=round(probg3,4)

# Click to paste the text onto the graph

text(4.5,0.05, paste("Area = ", probg3, sep=""))
```

## Task 4

Plotting the chi-square distributions.

```{r}
layout(matrix(1:4, nr=2,nc=2))

curve(dchisq(x,df=1),xlim=c(0,10),ylim=c(0,1),col="Red",lwd=2,
ylab="Chisq density", main="df=1")
curve(dchisq(x,df=2),xlim=c(0,10),ylim=c(0,0.5),col="Red",lwd=2,
ylab="Chisq density", main="df=2")
curve(dchisq(x,df=4),xlim=c(0,10),ylim=c(0,0.3),col="Red",lwd=2,
ylab="Chisq density", main="df=4")
curve(dchisq(x,df=20),xlim=c(0,40),ylim=c(0,0.1),col="Red",lwd=2,
ylab="Chisq density", main="df=20")
```

Plotting probabilities for the chi-square distributions.

```{r}
# First curve
curve(dchisq(x, df=2),xlim=c(0,10))

# Find the area between 2 and 4

# x values corresponding to the x - cords of points on the curve
xcurvec1=seq(2,4,length=1000)

# Y values corresponding t0 the x values
ycurvec1=dchisq(xcurvec1,df=2)

# Fill in the polygon with the given vertices
polygon(c(2,xcurvec1,4),c(0,ycurvec1,0),col="Red")

# Put in the text with the appropriate area

# Area
probc1=pchisq(4,df=2) - pchisq(2, df=2)
probc1=round(probc1,4)

# Click to paste the text onto the graph

text(3,0.05, paste("Area = ", probc1, sep=""))
```

```{r}
# Second curve
curve(dchisq(x, df=3),xlim=c(0,10))

# Find the area between 3 and 5

# x values corresponding to the x - cords of points on the curve
xcurvec2=seq(3,5,length=1000)

# Y values corresponding t0 the x values
ycurvec2=dchisq(xcurvec2,df=3)

# Fill in the polygon with the given vertices
polygon(c(3,xcurvec2,5),c(0,ycurvec2,0),col="Red")

# Put in the text with the appropriate area

# Area
probc2=pchisq(5,df=3) - pchisq(2, df=3)
probc2=round(probc2,4)

# Click to paste the text onto the graph

text(4,0.05, paste("Area = ", probc2, sep=""))
```

```{r}
# Third curve
curve(dchisq(x, df=20),xlim=c(0,40))

# Find the area between 10 and 21

# x values corresponding to the x - cords of points on the curve
xcurvec3=seq(10,21,length=1000)

# Y values corresponding t0 the x values
ycurvec3=dchisq(xcurvec3,df=20)

# Fill in the polygon with the given vertices
polygon(c(10,xcurvec3,21),c(0,ycurvec3,0),col="Red")

# Put in the text with the appropriate area

# Area
probc3=pchisq(21,df=20) - pchisq(10, df=20)
probc3=round(probc3,4)

# Click to paste the text onto the graph

text(15,0.02, paste("Area = ", probc1, sep=""))
```

## Task 5

Plotting the weibull distributions.

```{r}
#Setting up layout

layout(matrix(1:4, nr=2,nc=2))

#layout.show(4)

#Making the weibull distribution functions

curve(dweibull(x, shape=1,scale=1),xlim=c(0,25))
curve(dweibull(x, shape=2,scale=3),xlim=c(0,25))
curve(dweibull(x, shape=5,scale=8),xlim=c(0,25))
curve(dweibull(x, shape=13,scale=21),xlim=c(0,25))

```

Plotting probabilities for the weibull distributions.

```{r}
# First curve
curve(dweibull(x, shape=1,scale=5),xlim=c(0,25))

# Find the area between 2 and 8

# x values corresponding to the x - cords of points on the curve
xcurvew1=seq(2,8,length=1000)

# Y values corresponding t0 the x values
ycurvew1=dweibull(xcurvew1,shape=1,scale=5)

# Fill in the polygon with the given vertices
polygon(c(2,xcurvew1,8),c(0,ycurvew1,0),col="Red")

# Put in the text with the appropriate area

# Area
probw1=pweibull(8,shape=1,scale=5) - pweibull(2,shape=1,scale=5)
probw1=round(probw1,4)

# Click to paste the text onto the graph

text(5,0.05, paste("Area = ", probw1, sep=""))
```

```{r}
# Second curve
curve(dweibull(x, shape=11,scale=7),xlim=c(0,25))

# Find the area between 4 and 6

# x values corresponding to the x - cords of points on the curve
xcurvew2=seq(4,6,length=1000)

# Y values corresponding t0 the x values
ycurvew2=dweibull(xcurvew2,shape=11,scale=7)

# Fill in the polygon with the given vertices
polygon(c(4,xcurvew2,6),c(0,ycurvew2,0),col="Red")

# Put in the text with the appropriate area

# Area
probw2=pweibull(6,shape=11,scale=7) - pweibull(4,shape=11,scale=7)
probw2=round(probw2,4)

# Click to paste the text onto the graph

text(5,0.05, paste("Area = ", probw2, sep=""))
```

```{r}
# Third curve
curve(dweibull(x, shape=3,scale=11),xlim=c(0,25))

# Find the area between 13 and 20

# x values corresponding to the x - cords of points on the curve
xcurvew3=seq(13,20,length=1000)

# Y values corresponding t0 the x values
ycurvew3=dweibull(xcurvew3,shape=3,scale=11)

# Fill in the polygon with the given vertices
polygon(c(13,xcurvew3,20),c(0,ycurvew3,0),col="Red")

# Put in the text with the appropriate area

# Area
probw3=pweibull(20,shape=3,scale=11) - pweibull(13,shape=3,scale=11)
probw3=round(probw3,4)

# Click to paste the text onto the graph

text(16,0.03, paste("Area = ", probw3, sep=""))
```

## Task 6

Plotting the beta distributions.

```{r}
#Setting up layout

layout(matrix(1:4, nr=2,nc=2))

curve(dbeta(x, shape1=1, shape2=5))
curve(dbeta(x, shape1=5, shape2=1))
curve(dbeta(x, shape1=2, shape2=5))
curve(dbeta(x, shape1=6, shape2=8))
```

Plotting probabilities for the beta distributions.

```{r}
# First curve

curve(dbeta(x, shape1=1,shape2=7))

# Find the area between 0.4 and 1

# x values corresponding to the x - cords of points on the curve
xcurveb1=seq(0.4,1,length=1000)

# Y values corresponding t0 the x values
ycurveb1=dbeta(xcurveb1,shape1=1,shape2=7)

# Fill in the polygon with the given vertices
polygon(c(0.4,xcurveb1,1),c(0,ycurveb1,0),col="Red")

# Put in the text with the appropriate area

# Area
probb1=pbeta(1,shape1=1,shape2=7) - pbeta(0.4,shape1=1,shape2=7)
probb1=round(probb1,4)

# Click to paste the text onto the graph

text(0.5,2, paste("Area = ", probb1, sep=""))
```

```{r}
# Second curve

curve(dbeta(x, shape1=9,shape2=3))

# Find the area between 0.3 and 0.7

# x values corresponding to the x - cords of points on the curve
xcurveb2=seq(0.3,0.7,length=1000)

# Y values corresponding t0 the x values
ycurveb2=dbeta(xcurveb2,shape1=9,shape2=3)

# Fill in the polygon with the given vertices
polygon(c(0.3,xcurveb2,0.7),c(0,ycurveb2,0),col="Red")

# Put in the text with the appropriate area

# Area
probb2=pbeta(0.7,shape1=9,shape2=3) - pbeta(0.3,shape1=9,shape2=3)
probb2=round(probb2,4)

# Click to paste the text onto the graph

text(0.58,0.25, paste("Area = ", probb2, sep=""))
```

```{r}
# Third curve

curve(dbeta(x, shape1=2,shape2=3))

# Find the area between 0 and 0.5

# x values corresponding to the x - cords of points on the curve
xcurveb3=seq(0,0.5,length=1000)

# Y values corresponding t0 the x values
ycurveb3=dbeta(xcurveb3,shape1=2,shape2=3)

# Fill in the polygon with the given vertices
polygon(c(0,xcurveb3,0.5),c(0,ycurveb3,0),col="Red")

# Put in the text with the appropriate area

# Area
probb3=pbeta(0.5,shape1=2,shape2=3) - pbeta(0,shape1=2,shape2=3)
probb3=round(probb3,4)

# Click to paste the text onto the graph

text(0.25,0.7, paste("Area = ", probb3, sep=""))
```

## Task 7

Calling the function myncurve.

```{r}
MATH4753GROE002::myncurve(mu=10,sigma=5, a=6)
```



