---
title: 'Lab 12: MATH 4753'
author: "Dan Groeneveld"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
set.seed(55);x1=rnorm(30,mean=25,sd=5)
```

### H0: mu = 22

```{r}
t.test(x1, mu =22)
```

I reject the null hypothesis that mu=22 and accept the alternate hypothesis that mu is not equal to 22.

### H0: mu =23

```{r}
t.test(x1, mu =23)
```

I reject the null hypothesis that mu=23 and accept the alternate hypothesis that mu is not equal to 23.

### H0: mu =24

```{r}
t.test(x1, mu =24)
```

I accept the null hypothesis that 24 is a reasonable value for mu.

### H0: mu =25

```{r}
t.test(x1, mu =25)
```

I accept the null hypothesis that 25 is a reasonable value for mu.


### H0: mu =26

```{r}
t.test(x1, mu =26)
```

I accept the null hypothesis that 26 is a reasonable value for mu.

### Boxplot

```{r}
boxplot(x1, main="Sample x1")
ci=t.test(x1)$conf.int
ci
abline(h=c(ci,mean(x1)),col=c("Red","Red","Green"))
```

### Using pivotal statistic

```{r}
tcalc=(mean(x1)-24)/(sd(x1)/sqrt(30))
tcalc
```

```{r}
#### FUNCTION for Pvalues 
# Display P-value areas
mypvalue=function(t0,xmax=4,n=20, alpha=0.05){
#calculate alpha/2
va=round(pt(-t0,df=n-1),4)
pv=2*va

# plot the t dist
curve(dt(x,df=n-1),xlim=c(-xmax,xmax),ylab="T Density",xlab=expression(t),
main=substitute(paste("P-value=", pv, " alpha=", alpha)))


# set up points on the polygon to the right
xcurve=seq(t0,xmax,length=1000)
ycurve=dt(xcurve,df=n-1)

# set up points to the left
xlcurve=seq(-t0,-xmax,length=1000)
ylcurve=dt(xcurve,df=n-1)

# Shade in the polygon defined by the line segments
polygon(c(t0,xcurve,xmax),c(0,ycurve,0),col="green")
polygon(c(-t0,xlcurve,-xmax),c(0,ylcurve,0),col="green")

# make quantiles
q=qt(1-alpha/2,n-1)
abline( v=c(q,-q),lwd=2) # plot the cut off t value 
axis(3,c(q,-q),c(expression(abs(t[alpha/2])),expression(-abs(t[alpha/2]))))


# Annotation
text(0.5*(t0+xmax),max(ycurve),substitute(paste(area, "=",va)))
text(-0.5*(t0+xmax),max(ycurve),expression(area))

return(list(q=q,pvalue=pv))
}
```

```{r}
mypvalue(tcalc,n=30,alpha=0.05)
```

The rejection region is t is greater than 2.0452 or less than -2.0452.

The P-value that will determine if we will reject H0 or not is when p is less than alpha or in this case less than 0.05.

No, tcalc = 1.3263 is less than 2.0452 so it is not in the rejection region.

### Bootstrap intervals

```{r}
### bootstrap pvalues
bootpval<-function(x,conf.level=0.95,iter=3000,mu0=0, test="two"){
n=length(x)
y=x-mean(x)+mu0  # transform the data so that it is centered at the NULL
rs.mat<-c()    #rs.mat will become a resample matrix -- now it is an empty vector
xrs.mat<-c()
for(i in 1:iter){ # for loop - the loop will go around iter times
rs.mat<-cbind(rs.mat,sample(y,n,replace=TRUE)) #sampling from y cbind -- column bind -- binds the vectors together by columns
xrs.mat<-cbind(xrs.mat,sample(x,n,replace=TRUE)) #sampling from x cbind -- column bind -- binds the vectors together by columns

}

tstat<-function(z){ # The value of t when the NULL is assumed true (xbar-muo)/z/sqrt(n)
sqrt(n)*(mean(z)-mu0)/sd(z)
}

tcalc=tstat(x) # t for the data collected
ytstat=apply(rs.mat,2,tstat) # tstat of resampled y's, ytstat is a vector and will have iter values in it
xstat=apply(xrs.mat,2,mean)  # mean of resampled x's
alpha=1-conf.level # calculating alpha
ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
pvalue=ifelse(test=="two",length(ytstat[ytstat>abs(tcalc) | ytstat < -abs(tcalc)])/iter,
ifelse(test=="upper",length(ytstat[ytstat>tcalc])/iter,
length(ytstat[ytstat<xstat])/iter))

h=hist(ytstat,plot=FALSE)
mid=h$mid
if(test=="two"){
ncoll=length(mid[mid<= -abs(tcalc)])
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll-ncolr),rep("Green",ncolr))
}
if(test=="upper"){
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Gray",length(mid)-ncolr),rep("Green",ncolr))
}

if(test=="lower"){
ncoll=length(mid[mid<=  -abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll))
}
hist(ytstat,col=col,freq=FALSE,las=1,main="",xlab=expression(T[stat]))
#segments(ci[1],0,ci[2],0,lwd=2)
pround=round(pvalue,4)
title(substitute(paste(P[value],"=",pround)))
return(list(pvalue=pvalue,tcalc=tcalc,n=n,x=x,test=test,ci=ci))
}
```

```{r}
boot22=bootpval(x=x1,mu0=22,test="two")
boot22
```

```{r}
boot23=bootpval(x=x1,mu0=23,test="two")
boot23
```

```{r}
boot24=bootpval(x=x1,mu0=24,test="two")
boot24
```

```{r}
boot25=bootpval(x=x1,mu0=25,test="two")
boot25
```

```{r}
boot26=bootpval(x=x1,mu0=26,test="two")
boot26
```

These results match the results I found above. The null hypotheses that mu = 22 and 23 were rejected while the null hypotheses that mu = 24,25 and 26 were accepted as plausible. 

## Task 3

```{r}
set.seed(30);x=rnorm(15,mean=10,sd=7)   
set.seed(40);y=rnorm(20,mean=12,sd=4)
```

```{r}
var.test(x,y)
```

I conclude that the variances of these two samples being equal is not a likely scenario and reject the hypothesis that they are equal. 

As a result, I will assign var.equal the value of FALSE.

```{r}
t.test(y,x, mu = 0, var.equal = FALSE)
```

```{r}
t.test(y,x, mu=2,var.equal=FALSE)
```

From this, I conclude that the null hypothesis muy - mux = 0 as well as the null hypothesis muy - mux = 2 are both plausible for this data.

## Task 4

```{r}
set.seed(30);x=rnorm(15,mean=10,sd=4)   
set.seed(40);y=rnorm(20,mean=12,sd=4)

```

```{r}
var.test(x,y)
```

From this I conclude the variances being equal is a reasonable hypothesis for these samples and will use var.equal = TRUE in t.test().

```{r}
t.test(y,x, mu=0,var.equal=TRUE)
```

```{r}
t.test(y,x, mu=2,var.equal=TRUE)
```

From this, I conclude that the null hypothesis muy - mux = 2 is plausible given this data while I reject the null hypothesis muy - mux = 0.

## Task 5

```{r}
## Bootstrap interval for a two sample test
boot2pval<-function(x1,x2,conf.level=0.95,iter=3000,mudiff=0, test="two"){
n1=length(x1)
n2=length(x2)
y1=x1-mean(x1)+mean(c(x1,x2))  # transform the data so that it is centered at the NULL
y2=x2-mean(x2)+mean(c(x1,x2))
y1rs.mat<-c()    #rs.mat will be come a resample matrix -- now it is an empty vector
x1rs.mat<-c()
y2rs.mat<-c()
x2rs.mat<-c()
for(i in 1:iter){ # for loop - the loop will go around iter times
y1rs.mat<-cbind(y1rs.mat,sample(y1,n1,replace=TRUE)) #sampling from y cbind -- column bind -- binds the vectors together by columns
y2rs.mat<-cbind(y2rs.mat,sample(y2,n2,replace=TRUE))

}
x1rs.mat<-y1rs.mat+mean(x1)-mean(c(x1,x2))
x2rs.mat<-y2rs.mat+mean(x2)-mean(c(x1,x2))

xbar1=mean(x1)
xbar2=mean(x2)
sx1sq=var(x1)
sx2sq=var(x2)

tcalc=(xbar1-xbar2-mudiff)/sqrt(sx1sq/n1+sx2sq/n2)

sy1sq=apply(y1rs.mat,2,var)
sy2sq=apply(y2rs.mat,2,var) 
y1bar=apply(y1rs.mat,2,mean)
y2bar=apply(y2rs.mat,2,mean)

tstat=(y1bar-y2bar-mudiff)/sqrt(sy1sq/n1+sy2sq/n2)


alpha=1-conf.level # calculating alpha
#ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
pvalue=ifelse(test=="two",length(tstat[tstat>abs(tcalc) | tstat < -abs(tcalc)])/iter,
ifelse(test=="upper",length(tstat[tstat>tcalc])/iter,
length(ytstat[tstat<tcalc])/iter))

h=hist(tstat,plot=FALSE)
mid=h$mid
if(test=="two"){
ncoll=length(mid[mid<= -abs(tcalc)])
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll-ncolr),rep("Green",ncolr))
}
hist(tstat,col=col,freq=FALSE)
#segments(ci[1],0,ci[2],0,lwd=2)

return(list(pvalue=pvalue))
#return(list(pvalue=pvalue,tcalc=tcalc,n=n,x=x,test=test,ci=ci))
}
```

```{r}
set.seed(30);x=rnorm(15,mean=10,sd=7)   
set.seed(40);y=rnorm(20,mean=12,sd=4)
```

```{r}
boot2pval(x1=y, x2=x, mudiff = 0)
```

```{r}
boot2pval(x1=y, x2=x, mudiff = 2)
```



## Task 6

```{r}
set.seed(30);x=rnorm(15,mean=10,sd=4)   
set.seed(40);y=rnorm(20,mean=12,sd=4)
```

```{r}
boot2pval(x1=y, x2=x, mudiff = 0)
```

```{r}
boot2pval(x1=y, x2=x, mudiff = 2)
```

## Task 7

Line A calls the function t.test() on the sample x1 with the null hypothesis that the population mean = 23.

Line B indicates that this t-test is for a single sample as the only sample given when the function was called was x1.

Line C gives various results from the t-test: the value of t, t=2.3563, the degrees of freedom, df =29, and the p-value, p-value = 0.02543. From this, as the p-value is less than alpha = 0.05 (the default value in t.test()) we can reject the null hypothesis that mu = 23.

Line D states the alternative hypothesis, that the population mean is not equal to 23, which is accepted in this case.

Line E gives the percent confidence interval for this t.test(), 95%. This also means that alpha = 0.05 in this instance.

Line F gives the 95% confidence interval for the population mean, 23.30198 to 27.27320.

Line G gives the sample mean, 25.28759.

## Task 8

```{r}
set.seed(30);x=rnorm(15,mean=13,sd=4)   
set.seed(40);y=rnorm(20,mean=10,sd=4)
MATH4753GROE002::myboot2pval(x1 = x, x2 =y, mudiff=3)
```










