---
title: "Class Notes + Useful Techniques"
author: "Dan Groeneveld"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 7: Nov 4, 2020
## Calculating confidence intervals usting t.test()

## ATTENTION- this will be on upcoming exam!

### data:

```{r}
y1 <- rnorm(30, 25, 10)
y2 <- rnorm(40, 20, 10)
```

### using t.test() to calculate 86% confidence interval:

This is a confidence interval for mu1 - mu2: (actual value is 5)

```{r}
t.test(y1,y2,var.equal=TRUE, conf.level = .86)$conf.int
```

Easy shit babyyyyyyyyyy

## Paired Samples

Two samples that are connected, for example: two exam grades for the same student. The two different exam grades are dependent because each depends on the same student. Cannot take T-test of both exam scores because they are NOT independent.

However:
Exam2 - Exam1 is a meaningful statistic. In R: 

```{r}
#t.test(Exam1, Exam2, paired = TRUE,)$conf.int
```

or 

```{r}
#t.test(Exam1 - Exam2)$conf.int
```

You are trying to solve a problem where you are answering a question involving the difference (or similarity) between means in two different populations. Can show/compare relative sizes for both means, not their absolute size, but their relative size.
