---
title: "Lab 2"
author: "Dan Groeneveld"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```

## Task 2
```{r}
mpgdf = read.csv('EPAGAS.csv')

head(mpgdf)

```

## Task 3

```{r}
mpg = mpgdf$MPG
z=(mpg-mean(mpg))/sd(mpg)
mean(z)
sd(z)
mpg[abs(z)>=2 & abs(z)<=3]
mpg[abs(z)>3]

mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))
mypch = ifelse(abs(z)>3, 20,
        ifelse(abs(z)>=2 &abs(z)<=3,21, 22))
library(lattice)
dotplot(mpg, col=mycol)
```

## Task 4

```{r}
boxplot(mpg,xlab="MPG",col="Black",notch = TRUE, horizontal = TRUE, main = "MPG Boxplot")
```

Chebyshev's Rule predicts at least 75% of the data will fall within two standard deviations of the mean.

```{r}
length(mpg[abs(z)<2])/length(mpg)
```

Yes, Chebyshev agrees with the data as 96% falls within two standard deviations of the mean.

The Empirical Rule predicts 95% of the data will fall within two stand deviations of the mean.

The Empirical Rule corresponds very well to the data as it predicted the correct ratio of data to one percent.

The Empirical Rule is valid in this case because the data is roughly mound shaped and distributed symmetrically.


