---
title: "Lab 2"
author: "Dan Groeneveld"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```

## Task 2
```{r}
mpgdf = read.csv('EPAGAS.csv')

head(mpgdf)

```

## Task 3

Making the object mpg and transforming mpg to z:
```{r}
mpg = mpgdf$MPG
z=(mpg-mean(mpg))/sd(mpg)
```

Verifying results:
```{r}
mean(z)
sd(z)
```



The values of mpg that are possible outliers:
```{r}
mpg[abs(z)>=2 & abs(z)<=3]
```

The values of mpg that are defined as outliers:
```{r}
mpg[abs(z)>3]
```

```{r}
mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))
mypch = ifelse(abs(z)>3, 20,
        ifelse(abs(z)>=2 &abs(z)<=3,21, 22))
library(lattice)
dotplot(mpg, col=mycol, main = "MPG Dotplot")
```


## Task 4

```{r}
boxplot(mpg,xlab="MPG",col="Black",notch = TRUE, horizontal = TRUE, main = "MPG Boxplot")
```

Chebyshev's Rule stats that at least 1 - 1/(k^2) percent of the data will fall within k standard deviations of the mean.

Calculation: 1 - 1/(k^2) = 1- 1/(2^2) = 0.75

Chebyshev's Rule predicts at least 75% of the data will fall within two standard deviations of the mean.

```{r}
length(mpg[abs(z)<2])/length(mpg)
```

Yes, Chebyshev agrees with the data as 96% falls within two standard deviations of the mean.

The Empirical Rule predicts 95% of the data will fall within two stand deviations of the mean.

The Empirical Rule corresponds very well to the data as it predicted the correct ratio of data within one percent.

```{r}
hist(mpg, col = 1:8)
```


The Empirical Rule is valid in this case because the data is roughly mound shaped and distributed symmetrically.


