---
title: 'Lab 10: MATH 4753'
author: "Dan Groeneveld"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
#max likelihood function  
## For repeated sampling from same distribution                     
mymaxlik=function(lfun,x,param,...){
# how many param values are there?
np=length(param)
# outer -- notice the order, x then param
# this produces a matrix -- try outer(1:4,5:10,function(x,y) paste(x,y,sep=" "))   to understand
z=outer(x,param,lfun)
# z is a matrix where each x,param is replaced with the function evaluated at those values
y=apply(z,2,sum)

# y is a vector made up of the column sums
# Each y is the log lik for a new parameter value
plot(param,y,col="Blue",type="l",lwd=2,...)
# which gives the index for the value of y == max.
# there could be a max between two values of the parameter, therefore 2 indices
# the first max will take the larger indice
i=max(which(y==max(y)))
abline(v=param[i],lwd=2,col="Red")

# plots a nice point where the max lik is
points(param[i],y[i],pch=19,cex=1.5,col="Black")
axis(3,param[i],round(param[i],2))
#check slopes. If it is a max the slope shoud change sign from + to 
# We should get three + and two -vs
ifelse(i-3>=1 & i+2<=np, slope<-(y[(i-2):(i+2)]-y[(i-3):(i+1)])/(param[(i-2):(i+2)]-param[(i-3):(i+1)]),slope<-"NA")
return(list(i=i,parami=param[i],yi=y[i],slope=slope))
}
```

```{r}
outer(1:4,5:10,function(x,y) paste(x,y,sep=" "))
```


Line A produces are matrix of the values of a function (specified as lfun) which takes x and param (specified in the function) as input. This matrix is stored in the variable z.

Line B returns the value of y for which y is maximized. This value is then stored in the variable i.

The mathematical method:

$$ p(y_1, y_2,..., y_8 ) = p(y_1)p(y_2)...p(y_8)$$

$$= \binom{20}{y_1}p^{y_1}(1-p)^{20-y_1}\binom{20}{y_2}p^{y_2}(1-p)^{20-y_2}...\binom{20}{y_8}p^{y_8}(1-p)^{20-y_8}$$

$$ = \prod_{i=1}^{8}\binom{20}{y_i}p^{y_i}(1-p)^{20-y_i}$$
The R-code:

```{r}
#dbinom(y_1,20,p)*dbinom(y_2,20,p)*dbinom(y_3,20,p)*dbinom(y_4,20,p)*dbinom(y_5,20,p)*dbinom(y_6,20,p)*dbinom(y_7,20,p)*dbinom(y_8,20,p)
```

Estimation using graph:

```{r}
y = c(3,3,4,3,4,5,5,4)
logbin=function(x,param) log(dbinom(x,prob=param,size=20))
mymaxlik(lfun =logbin, x= y, param=seq(0,1,length=1000), main="Binomial, = n=20")
```

## Task 3

Graphical method:

```{r}
y = c(4,6,7,6,5) 
logpoiss=function(x,param) log(dpois(x,lambda=param)) 
mymaxlik(x=y,param=seq(0,20,length=1000),lfun=logpoiss,xlab=expression(lambda),main="Poisson",cex.main=2)
```

The algebraic equation:

$$ L(\lambda) = \frac{e^{-n\lambda}\sum_{i=1}^{n}Y_i}{\prod_{i=1}^{n}Y_i!}$$

Using my NRML:

```{r}
myNRML=function(x0,delta=0.001,llik,xrange,parameter="param"){
f=function(x) (llik(x+delta)-llik(x))/delta
fdash=function(x) (f(x+delta)-f(x))/delta
d=1000
i=0
x=c()
y=c()
x[1]=x0
y[1]=f(x[1])
while(d > delta & i<100){
i=i+1
x[i+1]=x[i]-f(x[i])/fdash(x[i])
y[i+1]=f(x[i+1])
d=abs(y[i+1])
}
layout(matrix(1:2,nr=1,nc=2,byrow=TRUE),width=c(1,2))
curve(llik(x), xlim=xrange,xlab=parameter,ylab="log Lik",main="Log Lik")
curve(f(x),xlim=xrange,xaxt="n", xlab=parameter,ylab="derivative",main=  "Newton-Raphson Algorithm \n on the derivative")
points(x,y,col="Red",pch=19,cex=1.5)
axis(1,x,round(x,2),las=2)
abline(h=0,col="Red")

segments(x[1:(i-1)],y[1:(i-1)],x[2:i],rep(0,i-1),col="Blue",lwd=2)
segments(x[2:i],rep(0,i-1),x[2:i],y[2:i],lwd=0.5,col="Green")

list(x=x,y=y)
}
```

```{r}
myNRML(x0 =1, delta = 0.000001,llik=function(x) log(dpois(4,x)*dpois(7,x)),xrange=c(0,20),parameter="lambda" )
```

The value of lambda hat given by myNRML is 5.49

## Task 4

```{r}
logbin2=function(theta){log(dbinom(2,prob=theta,size=6)) + log(dbinom(4,prob=theta,size=10))}
mymaxlikg=function(lfun="logbin2",theta) { # default log lik is a combination bin
nth=length(theta)  # nu. of valuse used in theta
thmat=matrix(theta,nr=nth,nc=1,byrow=TRUE) # Matrix of theta
z=apply(thmat,1,lfun) # z holds the log lik values
zmax=max(which(z==max(z)))  # finding the INDEX of the max lik
plot(theta,exp(z),type="l") # plot of lik
abline(v=theta[zmax],col="Blue")   #  verical line through max
axis(3,theta[zmax],round(theta[zmax],4))  # one tick on the third axis 
theta[zmax]   # theta corresponding to max lik
}

mymaxlikg(theta=seq(0,1,length=10000))
```

The graphical solution for p is 0.375.

## Task 5

Algebraic expression:

$$l(\theta_1,\theta_2) =\prod_{i=1}^{n}\binom{y_n}{y_i}\theta_1^{y_i}(1-\theta_1)^{20-y_i} \frac{e^{-n\theta_2}\sum_{i=1}^{n}Y_i}{\prod_{i=1}^{n}Y_i!}$$


```{r}
logbinpois=function(theta1,theta2) log(dbinom(4,size=20,prob=theta1)) + log(dpois(4,lambda=theta2))

maxlikg2=function(theta1,theta2,lfun="logbinpois",...){
n1=length(theta1)
n2=length(theta2)
z=outer(theta1,theta2,lfun)
contour(theta1,theta2,exp(z),...) # exp(z) gives the lik
maxl=max(exp(z))    # max lik
coord=which(exp(z)==maxl,arr.ind=TRUE)  # find the co-ords of the max
th1est=theta1[coord[1]] # mxlik estimate of theta1
th2est=theta2[coord[2]]
abline(v=th1est,h=th2est)
axis(3,th1est,round(th1est,2))
axis(4,th2est,round(th2est,2),las=1)
list(th1est=th1est,th2est=th2est)
}
maxlikg2(theta1=seq(0,1,length=1000),theta2=seq(0,10,length=1000),nlevels=20)
```

From the graph theta1 is approximately 0.2 and theta2 is is approximately 4.

## Task 6

```{r}

mymlnorm=function(x,mu,sig,...){  #x sample vector
nmu=length(mu) # number of values in mu
nsig=length(sig)
n=length(x) # sample size
zz=c()    ## initialize a new vector
lfun=function(x,m,p) log(dnorm(x,mean=m,sd=p))   # log lik for normal
for(j in 1:nsig){
z=outer(x,mu,lfun,p=sig[j]) # z a matrix 
# col 1 of z contains lfun evaluated at each x with first value of mu, 
# col2 each x with 2nd value of m 
# all with sig=sig[j]
y=apply(z,2,sum)
# y is a vector filled with log lik values, 
# each with a difft mu and all with the same sig[j]
zz=cbind(zz,y)
## zz is the matrix with each column containing log L values, rows difft mu, cols difft sigmas 
}
maxl=max(exp(zz))
coord=which(exp(zz)==maxl,arr.ind=TRUE)
maxlsig=apply(zz,1,max)
contour(mu,sig,exp(zz),las=3,xlab=expression(mu),ylab=expression(sigma),axes=TRUE,
main=expression(paste("L(",mu,",",sigma,")",sep="")),...)
mlx=round(mean(x),2)  # theoretical
mly=round(sqrt((n-1)/n)*sd(x),2)
#axis(1,at=c(0:20,mlx),labels=sort(c(0:20,mlx)))
#axis(2,at=c(0:20,mly),labels=TRUE)
abline(v=mean(x),lwd=2,col="Green")
abline(h=sqrt((n-1)/n)*sd(x),lwd=2,col="Red")

# Now find the estimates from the co-ords
muest=mu[coord[1]]
sigest=sig[coord[2]]

abline(v=muest, h=sigest)
return(list(x=x,coord=coord,maxl=max, muest=muest, sigest = sigest))
}

mymlnorm(x=c(10,12,13,15,12,11,10),mu=seq(10,15,length=1000),sig=seq(0.1,4,length=1000),lwd=2,labcex=1)
```

The graphical estimate for the mean is 11.8569 and 1.6420 for the standard deviation.

## Task 7

```{r}
sam= rbeta(30,shape1=3,shape2=4)
mymlbeta=function(x,alpha,beta,...){  #x sample vector
na=length(alpha) # number of values in alpha
nb=length(beta)
n=length(x) # sample size
zz=c()    ## initialize a new vector
lfun=function(x,a,b) log(dbeta(x,shape1=a,shape2=b))   # log lik for beta
for(j in 1:nb){
z=outer(x,alpha,lfun,b=beta[j]) # z a matrix 
# col 1 of z contains lfun evaluated at each x with first value of alpha, 
# col2 each x with 2nd value of a 
# all with b=beta[j]
y=apply(z,2,sum)
# y is a vector filled with log lik values, 
# each with a difft alpha and all with the same sig[j]
zz=cbind(zz,y)
## zz is the matrix with each column containing log L values, rows difft alpha, cols difft betas 
}
maxl=max(exp(zz))    # max lik
coord=which(exp(zz)==maxl,arr.ind=TRUE)  # find the co-ords of the max
aest=alpha[coord[1]] # mxlik estimate of alpha
best=beta[coord[2]]
contour(alpha,beta,exp(zz),las=3,xlab=expression(alpha),ylab=expression(beta),axes=TRUE,
main=expression(paste("L(",alpha,",",beta,")",sep="")),...)

abline(v=aest, h=best)
points(aest,best,pch=19)
axis(4,best,round(best,2),col="Red")
axis(3,aest,round(aest,2),col="Red")
return(list(x=x,coord=coord,maxl=maxl,maxalpha=aest,maxbeta=best))
}

mymlbeta(x=sam,alpha=seq(1,4,length=100),beta=seq(2,6,length=100),lwd=2,labcex=1)
```

```{r}
layout(matrix(1:12,nr=3,nc=4,byrow=TRUE))
z=c()
a=3
b=4
sam= rbeta(30,shape1=a,shape2=b)
nsam=length(sam)
for(i in 1:12){
w=mymlbeta(x=sample(sam,nsam,replace=TRUE),alpha=seq(0.1,20,length=100),beta=seq(0.1,20,length=100),lwd=2,labcex=1,col="steelblue")
points(a,b,col="Red",pch=19)
abline(v=a,h=b,col="Red")
z<-rbind(z,c(w$maxalpha,w$maxbeta))
 }
colnames(z)=c("alpha","beta")
z
dz=apply(z-c(a,b),1,function(x) sqrt(sum(x^2)) )
plot(dz)
which(dz>4)
```

## Task 8

Adding the function my package.

Proving it's been added:

```{r}
MATH4753GROE002::mymaxlikdg(x=c(9,9,1,9,9,9),param=seq(0,1,length=1000),lfun=logbin,xlab=expression(pi),main="Binomial",cex.main=2)
```




