---
title: "Final Exam"
author: "Dan Groeneveld"
date: "12/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
qt(1-.02/2,24)
```
```{r}
invq = read.csv("INVQUAD.csv")
xe <- invq[invq$PLANT=="Electric","INV.QUAD"]
xg <- invq[invq$PLANT=="Gas","INV.QUAD"]
vare = var(xe)
varg = var(xg)
vare/varg
```
```{r}
var.test(xe ,xg, conf.level=.95)
```

```{r}
t.test(xe, xg ,conf.level=.95, var.equal = TRUE)
```


#Filling in values!!

y = c(1,2,3) # sample of n = 3 from normal population
mu = mean(y)
sig = sd(y)
tcalc = (mu -1)/(sig/sqrt(3)) # the one is the null hypothesis
tcalc

2*(1-pt(tcalc, df = n -1)) pvalue!!

# T-test

```{r}

```


# SLR w/ S20x

```{r}
kwhrs = read.csv("KWHRS.csv")
x = kwhrs$SIZE
y = kwhrs$USAGE
ylm <- lm(y~x)
summary(ylm)
```

```{r}
library(s20x)
ciReg(ylm)
```

ylm <- lm(y~x)
summary(ylm)
confint(ylm)


library(s20x)
trendscatter(CH4~D,f=0.5, data=aic)
aic.lm = lm(CH4~D, data = aic)
with(aic, 
plot(CH4~D,bg="Blue",pch=21,ylim=c(0.9*min(CH4),1.1*max(CH4)),xlim=c(0.9*min(D),1.1*max(D)),main=("Trendline for data"))
)
abline(aic.lm)

```{r}

```


# Q-LM

quad.lm=lm(Height~BHDiameter + I(BHDiameter^2),data=spruce.df)
summary(quad.lm)



```{r}
qlm = lm(y~x + I(x^2))
summary(qlm)
```

```{r}
1 - pnorm(11, 10, 4)
pnorm(20,20,10) - pnorm(10,20,10)
```



```{r}

1 - pnorm(1-qnorm(0.7,0,1), 0, 1)
```




# calculating probabilities

tab1 <- with(mtbe, table(WellClass, Aquifier))
addmargins(tab1)

Bayes function:

mybayes = function( ppgu, pngnu,pu=0.005) # pu = prob user, ppgu = prob pos g user (sensitivity), pngnu = prob neg g not user
{

pnu=1-pu # pu is the prior, pnu = prob not user
pngu = 1-ppgu # pngu = prob neg given user
ppgnu = 1-pngnu # ppgnu = prob pos given not user
pp = pu*ppgu + pnu*ppgnu # pp = prob pos = evidence
pugp = pu*ppgu/pp # pugp = prob user g pos
pnugp=1-pugp # pnugp = prob not user g pos

list("post" = pugp,pnugp=pnugp, prior=pu, sensitivity = ppgu, specificity = pngnu)
}

mybayes(ppgu = 0.91, pngnu = 0.82, pu = 5/100)

